A weighted sampling algorithm for the design of RNA sequences with targeted secondary structure and nucleotides distribution

### 2013-03-12 ISMB/results_ISMB.tex

%Figure~\ref{fig:nb_sols_entropy} shows the number of correct solutions as their average entropy and positions inside base pairs entropy, since \ourprog  only influences the distribution of nucleotides inside stacked base pairs. The most constraining factor for generating valid solutions seems to be  low \GCContent and the percentage of free nucleotides.

@@

\begin{figure*}[ht!]
	\centering
	\includegraphics[width=\textwidth]{Figures/nb_sols_entropy}
	\caption{Number of solutions generated by \ourprog and their average sequence identity. The curves are parameterized by the {\GCContent} of the sequences. The data are plotted vs. the length of the target (left), density of stacked base pairs (centre) and number of unpaired nucleotides  in loops (right).}
	\label{fig:nb_sols_entropy}
\end{figure*}

@@

\highlight{
\subsection{Global sampling vs Local search vs Glocal approach}
To conclude this study, we estimate the impact of the design methodology on the performances. More precisely, we aim to determine the merits of a global sampling approach (\ourprog), compared to a glocal procedure (\ourprog + \RNAinverse) and a local search methodology (\RNASSD). To our knowledge, \RNASSD is the only other software that implements an explicit control of the \GCContent. We focus our analysis on the design of sequences with low {\GCContent}s (30\% and less) as they are almost impossible to design with classical software (See Figure~\ref{fig:gcdist}).

Our results (See Fig. \ref{fig:rnainverse} in Supplementary Material) indicate that a glocal approach produces sequences that are consistently more diverse than those obtained with both {\em pure} methods alone. Furthermore, while \ourprog outperforms \RNAinverse for extreme pressures on the \GCContent, \RNAinverse produces more entropic sequences than \ourprog for the central \GCContent regime ($30\%-70\%$).

It should be noted, however, than \RNAinverse produces sequences whose \GCContent is extremely concentrated (see Fig.), thereby requiring a large number of runs before producing a suitable \GCContent sequence. Therefore, a rejection algorithm based on \RNAinverse alone should not be considered a realistic solution for the design with constrained \GCContent. 
}

\begin{figure*}[ht!]
  \centering
  \includegraphics[width=\textwidth]{Figures/seq_identity_10_30}
  \caption{Sequence identity of \ourprog and \RNAinverse for $10$ and $30\%$ of \GC.}
\end{figure*}

>>>>

  Figure~\ref{fig:nb_sols_entropy} shows the number of correct
solutions \highlight{as their average sequence identity. Since \ourprog
only influences the distribution of nucleotides inside stacked base pairs, 
we consider the \GCContent only inside those base pairs.}
  %entropy and positions inside base pairs entropy, since 
%\ourprog  only influences the distribution of nucleotides inside 
%stacked base pairs. 
The most constraining factor for generating valid
 solutions seems to be  low \GCContent and the percentage of free nucleotides.

%Our method is able to generate 
%Figure~\ref{fig:nb_sols_entropy} shows the number of correct solutions as their average entropy and positions inside base pairs entropy, since \ourprog  only influences the distribution of nucleotides inside stacked base pairs. The most constraining factor for generating valid solutions seems to be  low \GCContent and the percentage of free nucleotides.

@@

\begin{figure*}[ht!]
	\centering
	\includegraphics[width=\textwidth]{Figures/nb_sols_entropy}
	\caption{Number of solutions generated with \ourprog+\RNAinverse. and their average sequence identity. 
	The second row is the sequence identity for a $10\%$\GC and the last for $30\%$\GC inside the base pairs. }
%	\caption{Number of solutions generated by \ourprog and their average sequence identity. The curves are parameterized by the {\GCContent} of the sequences. The data are plotted vs. the length of the target (left), density of stacked base pairs (centre) and number of unpaired nucleotides  in loops (right).}
	\label{fig:nb_sols_entropy}
\end{figure*}

@@

%<<<<<<< HEAD

%\begin{figure*}[ht!]
%  \centering
%  \includegraphics[width=\textwidth]{Figures/seq_identity_10_30}
%  \caption{Sequence identity of \ourprog and \RNAinverse for $10$ and $30\%$ of \GC.}
%\end{figure*}

%\subsection{Global sampling vs Local search vs Glocal approach}

%Here, we are interested in estimating the impact of the design methodology on the performances. More precisely, we aim to determine the merits of a global sampling approach (\ourprog), compared to a glocal procedure (\ourprog + \RNAinverse) and local search methodologies (\RNAinverse, \INFORNA, \NUPACK and \frankenstein). 
%=======
\highlight{
\subsection{Global sampling vs Local search vs Glocal approach}
To conclude this study, we estimate the impact of the design methodology on the performances. More precisely, we aim to determine the merits of a global sampling approach (\ourprog), compared to a glocal procedure (\ourprog + \RNAinverse) and a local search methodology (\RNASSD). To our knowledge, \RNASSD is the only other software that implements an explicit control of the \GCContent. We focus our analysis on the design of sequences with low {\GCContent}s (30\% and less) as they are almost impossible to design with classical software (See Figure~\ref{fig:gcdist}).
%>>>>>>> af029e5428ff6e8980ef29cd0002a6fd72501bd9

Our results (See Fig. \ref{fig:no_gc_controle} in Supplementary Material) indicate that a glocal approach produces sequences that are consistently more diverse than those obtained with both {\em pure} methods alone. Furthermore, while \ourprog outperforms \RNAinverse for extreme pressures on the \GCContent, \RNAinverse produces sequences with a lower sequence identity than \ourprog for the central \GCContent regime ($30\%-70\%$).

%<<<<<<< HEAD
%It should be noted, however, that \RNAinverse, \INFORNA, \NUPACK  and \frankenstein produce sequences whose \GCContent is extremely concentrated (see Fig.~\ref{fig:gcdist}), thereby requiring a large number of runs before producing a suitable \GCContent sequence. Therefore, a rejection algorithm based on \RNAinverse alone should not be considered a realistic solution for the design with constrained \GCContent. 
%=======
It should be noted, however, than \RNAinverse produces sequences whose \GCContent is extremely concentrated (see Fig.~\ref{fig:gcdist}), thereby requiring a large number of runs before producing a suitable \GCContent sequence. Therefore, a rejection algorithm based on \RNAinverse alone should not be considered a realistic solution for the design with constrained \GCContent. 
}
%>>>>>>> af029e5428ff6e8980ef29cd0002a6fd72501bd9

%\begin{figure*}[ht!]
%  \centering
%  \includegraphics[width=\textwidth]{Figures/seq_identity_10_30}
%  \caption{Sequence identity of \ourprog and \RNAinverse for $10$ and $30\%$ of \GC.}
%\end{figure*}
